---
title: "Generating Multivariate Long-range-dependent Time series"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: https://github.com/ygeunkim/sim-lrd/blob/master/index.Rmd
runtime: shiny
---

```{r global, message=FALSE, include=FALSE}
# shiny------------------------
library(shiny)
library(flexdashboard)
# wrangling data and plot------
library(tidyverse)
library(plotly)
library(forecast)
# time series------------------
library(multiwave)
```

Sidebar {.sidebar}
======================================================================

```{r}
numericInput("seed", label = "Random seed:", value = 1, min = 1, step = 1)
sliderInput("N", label = "Time Series Size:", min = 100, max = 5000, value = 200, step = 100)
sliderInput("m", label =  "Time series Dimension:", min = 2, max = 20, value = 2, step = 1)
# sliderInput("ndim", label =  "Block Number of cov_matrix:", min = 1, max = 10, value = 1, step = 1)
sliderInput("burnin", label =  "Burn-in:", min = 1000, max = 10000, value = 2000, step = 1000)
sliderInput("lag.max", label = "ACF Lag until:", min = 10, max = 1000, value = 100, step = 1)
```

See **How to** menu why this generated dataset is long-range dependent process.

Download `cor.csv` and change value of correlation matrix for I(d) process if you want:

```{r}
# correlation upload----------------------
cov_data <- reactive({
  COV <- diag(input$m)
  colnames(COV) <- paste0("x", 1:ncol(COV))
  COV %>% 
    as.data.frame()
})
output$covdt <- downloadHandler(
  filename = "cor.csv",
  content = function(file) {
    cov_data() %>% 
      write_csv(file = file)
  }
)
downloadLink("covdt", label = "Change correlation matrix")
fileInput(
  "covfile", 
  label = "Download and reupload the file", 
  multiple = FALSE, 
  accept = ".csv",
  buttonLabel = "Browse/drag file",
  placeholder = "No file selected"
)
```

You can download the simulation result, `lrd.csv`:

```{r}
# dataset---------------------------------
lrd_data <- reactive({
  # input---------------------------------
  if (is.null(input$covfile)) {
    COV <- diag(input$m)
  } else {
    COV <- read.csv(input$covfile$datapath, header = TRUE, stringsAsFactors = FALSE) %>% as.matrix()
  }
  # COV <- kronecker(diag(input$ndim), COV)
  # LRD parameter---------------------------
  d <- runif(input$m, .1, .45)
  d <- floor(100 * d) / 100
  # generate------------------------------
  set.seed(input$seed)
  fivarma(
    N = input$N, d = d, cov_matrix = COV,
    VAR = NULL, VMA = NULL,
    skip = input$burnin
  )$x %>% 
    as.data.frame() %>% 
    tibble() %>% 
    rename_all(~str_replace_all(., pattern = "V", replacement = "x"))
})
# download button-------------------------
output$data <- downloadHandler(
  filename = "lrd.csv",
  content = function(file) {
    lrd_data() %>% 
      write_csv(file = file)
  }
)
downloadLink("data", label = "Download Vector LRD Series")
```

```{r}
selectInput(
  "viridis", label = "Viridis color scale of ggplot:", 
  choices = c("default ggplot", "magma", "inferno", "plasma", "viridis (default viridis)", "cividis"),
  selected = "default ggplot",
  multiple = FALSE
)
```

The original code courtesy of professor [Changryong Baek](https://sites.google.com/view/crbaek).

LRD Simulation
======================================================================

Row
-----------------------------------------------------------------------

### Time plot

```{r}
# ggplot----------------------------------
gg_lrd <- reactive({
  p <- 
    lrd_data() %>% 
    mutate(time_point = 1:n()) %>% 
    pivot_longer(-time_point, names_to = "variable", values_to = "values") %>% 
    ggplot(aes(x = time_point)) +
    geom_path(aes(y = values, colour = variable), alpha = .7, show.legend = FALSE) +
    labs(
      x = "Time",
      y = element_blank()
    ) +
    theme_minimal()
  if (input$viridis != "default ggplot") {
    p <- 
      p +
      scale_colour_viridis_d(option = input$viridis)
  }
  p
})
```

```{r}
output$timeplot <- renderPlotly({
  ggplotly(gg_lrd())
})
plotlyOutput("timeplot")
```

Row
-----------------------------------------------------------------------

### Spectral density (Periodogram)

```{r}
spectral <- reactive({
  spec <- 
    lrd_data() %>% 
    spectrum(plot = FALSE) %>% 
    .$spec %>% 
    as.data.frame()
  colnames(spec) <- paste0("x", 1:ncol(spec))
  p <- 
    spec %>% 
    mutate(time_point = 1:n()) %>% 
    pivot_longer(-time_point, names_to = "variable", values_to = "values") %>% 
    mutate(values = log(values)) %>% 
    ggplot(aes(x = time_point)) +
    geom_path(aes(y = values, colour = variable), alpha = .7, show.legend = FALSE) +
    labs(
      x = "log(Frequency)",
      y = "log(Periodogram)"
    ) +
    scale_x_continuous(trans = scales::log_trans()) +
    theme_minimal()
  if (input$viridis != "default ggplot") {
    p <- 
      p +
      scale_colour_viridis_d(option = input$viridis)
  }
  p
})
```

```{r}
output$periodogram <- renderPlotly({
  ggplotly(spectral())
})
plotlyOutput("periodogram")
```

Autocorrelation Function
======================================================================

```{r}
output$ccf <- renderPlotly({
  ggacf <- 
    ggAcf(lrd_data() %>% as.matrix(), lag.max = input$lag.max) +
    theme(axis.title = element_blank()) +
    labs(title = element_blank())
  ggplotly(ggacf)
})
plotlyOutput("ccf")
```


How to
======================================================================

### Fractionally integrated noise process

We generate (multivariate) long-range-dependent time series using (vector) $I(d)$ process (i.e. $FARIMA(0, d, 0)$).
Here, $d \in (0, 0.5)$ gives LRD series.

We choose $d$ randomly for each variable.

### $FARIMA(p, d, q)$

In case of 1-dim,

$$\phi(B) (1 - B)^d X_t = \theta(B) Z_t$$

When m-dim, i.e. 

$$\Phi(B) \mathbf{U}_t = \Theta(B) \mathbf{Z}_t$$

where

- VAR(p) term: $\Phi(B) = I_m - \Phi_1 B - \cdots - \Phi_p B^p$
- VMA(q) term: $\Theta(B) = I_m + \Theta_1 B + \cdots + \Theta_q B^q$
- Vector I(d) term: $\mathbf{U}_t = diag((I - B)^d)$

Note that $\mathbf{Z_t}$ is (vector) white noise of mean zero.
By default, its variance matrix is identity and we can change the value at the right `r emo::ji("smile")`

